---
title: "Final project"
author: "Jiaxuan Li"
date: "11/16/2020"
output: html_document
---
Summary: Because I currently don't have data in hand, neither do our lab members. Thus, I downloaded the COVID-19 data from WHO (https://covid19.who.int/table).This project is focuesed on the COVID-19 data in different countries from January to November in order to explore the pandemic situation. ALthough a lot of online sources had already done a lot of analysis of the COVID-19 data and they made good visualization plots. However, it's a good practice for me how to manipulate data, clean data, and mutate data no matter what data is.

In this project, I firstly ploted the COVID-19 data for China and the United States from Jan to Nov using simple line plot, boxplot and facet as well. Secondly, I 



```{r setup, include=FALSE}
##install packages and load in libraries
library(ggplot2)
install.packages("lubridate")
library(lubridate)
library(dplyr)
library(ggpubr)
library(viridis)
install.packages("sf")
library(sf)

```



1. Load in the dataset, which is downloaded from WHO(). Take a glimpse at the dataset.
```{r}
getwd()

##load in data
 data = read.csv("WHO-COVID-19-global-data.csv")
 
 ##take a look at the data
 str(data)
 glimpse(data)
 head(data)
 summary(data)
 class(data$Country)
 data$Country=as.factor(data$Country)
 class(data$Date_reported)
 class(data$Country_code)
 
```



2. Change the date settings firat since it's too intense if plot the data on every day, whcih will be used later. 

```{r}
##install "lubridate" package (https://rawgit.com/rstudio/cheatsheets/master/lubridate.pdf), which is specificlly designed for maniplating date related data
install.packages("lubridate")
library(lubridate)


##change the class of Date_reported column from "character" to "Date" class type
class(data$Date_reported)
data$Date_reported = as.Date(data$Date_reported)
class(data$Date_reported)

##change the breaks of the dates, whcih will be used later for plotting
date_breaks <- seq(as.Date("2020-01-03"), as.Date("2020-11-14"), by = "1 month" )

```


3. Plot the change of COVID-19 from Jan to Dec in China
```{r}
##filter function is in the dplyr package
library(dplyr)

##subset the covid-19 data for China
data_china = filter(data, data$Country_code == "CN")
dim(data_china)
head(data_china)

##Make the first line plot! Plot the COVID-19 data in China from January to December.
ggplot(data_china, aes(x=Date_reported, y=Cumulative_cases)) + 
  geom_point() + labs(title="Cumulative_cases in China") 

##Plot the Cumulative_cases in China from January to December.
gp_china_1 = ggplot(data_china, aes(x=Date_reported, y=Cumulative_cases)) + 
  geom_point(color='lightgreen', size = 0.5) + labs(title="Cumulative_cases in China") + 
  geom_line()+
  scale_x_date(breaks = date_breaks) +       
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2 , hjust=0.2))
gp_china_1

##Plot the new_cases in China from January to December.
gp_china_2 = ggplot(data_china, aes(x=Date_reported, y=New_cases)) +
  geom_point(color='lightblue', size = 0.5) + labs(title="New_cases in China") + 
  geom_line()+
  scale_x_date(breaks = date_breaks) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust=0.2))
gp_china_2


##Plot the new deaths in China from January to December.
gp_china_3 = ggplot(data_china, aes(x=Date_reported, y=New_deaths)) + 
  geom_point(color='lightblue', size = 0.5) + labs(title="New deaths in China") + 
  geom_line()+
  scale_x_date(breaks = date_breaks) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust=0.2))
gp_china_3

##Plot the cumulative deaths in China from January to December.
gp_china_4 = ggplot(data_china, aes(x=Date_reported, y=Cumulative_deaths)) + 
  geom_point(color='lightgreen', size = 0.5) + labs(title="Cumulative_deaths in China")  +
  geom_line()+
  scale_x_date(breaks = date_breaks) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust=0.2))
gp_china_4

```


```{r}
##Make a facet (https://www.datanovia.com/en/lessons/combine-multiple-ggplots-into-a-figure/) for the COVID-19 data in China
library(ggpubr)
ggarrange(gp_china_1, gp_china_2, gp_china_3, gp_china_4, ncol = 2, nrow = 2) + geom_jitter()
```


```{r}
###Make barplots for data in China; overlay a line plot with a box plot
bp_china1 = ggplot(data_china, aes(x=Date_reported, y=Cumulative_cases)) + 
  geom_bar(stat = "identity", color = "lightblue") + labs(title="Cumulative_cases in China") + 
  geom_line(aes(x=Date_reported, y=New_cases), color = "red") 
bp_china1

bp_china2 = ggplot(data_china, aes(x=Date_reported, y=Cumulative_deaths)) + 
  geom_bar(stat = "identity", color = "lightblue") + labs(title="Cumulative_deaths in China") + 
  geom_line(aes(x=Date_reported, y=New_deaths), color = "red") + legend (legend = "New deaths in China", col = red, cex = 0.8)
bp_china2 

##make a facet for the boxplots for the COVID-19 data in China
ggarrange(bp_china1, bp_china2, ncol = 2, nrow = 1) 
```


4. Plot the COVID-19 data in US from Jan to Dec

```{r}
##Subset the covid-19 data for US
data_US = filter(data, data$Country_code == "US")

##Make a simple line plot for the Cumulative_cases in the US
ggplot(data_US, aes(x=Date_reported, y=Cumulative_cases)) + 
  geom_point(color = "lightblue") + 
  geom_line()
  labs(title="Cumulative_cases in US") 

##Make a simple line plot for the Cumulative_cases in the US using a differnt x scale.
ggplot(data_US, aes(x=Date_reported, y=Cumulative_cases)) + 
  geom_point() + labs(title="Cumulative_cases in US") + 
  scale_x_date(breaks = date_breaks) +  
  theme(axis.text.x = element_text(angle = 45, vjust = 0, hjust=0))

##Make a facet plot for the covid-19 data in the US
##Build up a function and use lapply function to plot multiple plots at the same time
cases = c("Cumulative_cases", "New_cases", "Cumulative_deaths", "New_deaths")
plot_names = c ("gp_US1", "gp_US2", "gp_US3", "gp_US4")

plotmultiplelines = function (colnames){
  ggplot(data_US, aes_(x = ~Date_reported, y = as.name(colnames))) +
    geom_point()+ facet_wrap(~ variable, scales = "free_y") 
}


lapply(names(data_US[c(5:8)]), plotmultiplelines)



ggarrange(gp_china_1, gp_china_2, gp_china_3, gp_china_4, ncol = 2, nrow = 2) + geom_jitter()



```



5. Make heatmaps for several selected countries. 

```{r}
##Make a heatmap for all the countries in the original dataset
library(viridis)
ggplot(data, aes(x = Country, y =Date_reported, fill = Cumulative_cases))+geom_tile()

library(dplyr)
##The x axis is too intense, thus I will choose 5 countries randomly from the dataset and select them according to the country code.
selected_country = filter(data, Country_code == c("CN","US","GB","FR","CA"))
head(selected_country)

###Make the heatmap to illustrate the Cumulative_cases in 5 selected countries.
ggplot(selected_country, aes(y = Date_reported , x = Country, fill = Cumulative_cases))+
  scale_y_date(breaks = date_breaks) +  geom_tile()  + scale_fill_viridis_c() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.2, hjust=0.2))

##Here is another heatmap for the Cumulative_cases in the US.
ggplot(data_US, aes(y = Date_reported , x = Country, fill = Cumulative_cases))+
  scale_y_date(breaks = date_breaks) +  geom_tile()  + scale_fill_viridis_c() 
```
Brief Summary: 


6. Plot a map for several selected countries.
```{r}
install.packages("sf")
library(sf)
ggplot(data) + 
    geom_sf()


```





7. Make an interactive plot

```{r}

```
